Intensity Modulated Radiation Therapy using Tomotherapy

### 2014-12-08 BME 560/Paper.tex
Stuff added

  The rest of this paper is 
  organized as follows. Section \ref{history} talks 
  about the history of tomotherapy. Section \ref{methods} has 3 subsections: 
  \ref{construction} talks about the construction of the tomotherapy unit, 
  \ref{radiation} discusses the radiation characteristics of helical 
  tomotherapy, and \ref{comparisions} brielfy examines the differences between 
  this techique and other conformal therapy techniques. Section \ref{clinical} 
  is a collection of some clinical results where tomotherapy has been, and has 
  not been, effective. Finally, section \ref{conclusions} presents concluding 
  remarks on the the topic.

@@

  \subsection{Construction}
  \begin{figure}
    \centering
    \subfloat[\label{fig4a}][A simplified view of the Helical Tomotherapy 
    treatment head system in use today. The setup consists of a 6 MV linac to 
    produce an x-ray beam which is collimated down to a fan-beam using 2 
    moveable jaws (denoted as `y-jaws' in the image)]{
      \includegraphics[width=\linewidth]{fig4}}\\
    \subfloat[][]{
      \includegraphics[width=\linewidth]{fig5}}

  \end{figure}
  \label{construction}
  \subsubsection{A note on Serial Tomotherapy}
  \subsection{Radiation Characteristics}
  \label{radiation}
  \subsection{Comparisions between IMRT techniques}
  \label{comparisions}
    \begin{table*}[htbp!]
      \centering
      \caption{A comparision between standard and non-standard IMRT delivery 
      systems (Ref. \cite{Fenwick2006})}
      \small{
      \begin{tabulary}{\linewidth}{LLLLLLL}
        \toprule[0.04cm]

        \textbf{System} & \textbf{Gantry} & \textbf{ Beam Geometry} & 
        \textbf{Collimated} & \textbf{Modulated} & \textbf{Imaging} & 
        \textbf{Gating
          Potential} \\\toprule[0.04cm]

@@

      \label{tab:compare}%
    \end{table*}%
  Tomotherapy is different from other standard IMRT techiniques such as 
  conformal radiotherapy (CRT) techniques and non-standard IMRT techniques like 
  Cyber knife.   Standard CRT techniques generate highly conformal radiotherapy 
  using radiation fields generated in a few fixed-angle positions. These are 
  generated by linear accelerators (aka. LINAC) which are then modulated by 
  conventional leaf collimators operated in either dynamic mode or the standard 
  step-and-shoot mode. CyberKnife is a modern radiotherapy system developed by 
  Accuray, Inc., primarily for radiosurgery. It makes use of a highly precise 
  robotic arm to delivery the doses. The arm is computer driven to achieve very 
  high precision, it is mounted with a compact 6-MV LINAC whose orientation is 
  determined by 6 different mechanical subsystems. This robotic arm can be 
  moved with a much greater generality than most conventional sytsems. Table 
  \ref{tab:compare} highlights the key differences in various systems. For the 
  sake the completeness, the authors of Ref. \cite{Fenwick2006} have 
  dicotomized the conventional system into 2 types, with and without a 
  multileaf collimator, and the tomotherapy into serial and parallel 
  tomotherapy.
  \section{Clinical Results}
  \label{clinical}
  \subsection{Total Body Irradiation (TBI)}
  \subsection{Whole brain helical Tomotherapy}
  \subsection{Inoperable Lung Cancer}
  \subsection{Rectal Cancer}
  \section{Conclusion}
  \label{conclusions}
  \bibliography{Paper}{}
  \bibliographystyle{IEEEtran}

>>>>

  The rest of this paper is 
  organized as follows. Section \ref{history} talks 
  about the history of tomotherapy. Section \ref{methods} has 3 subsections: 
  \ref{construction} talks about the construction of the tomotherapy unit, 
  \ref{radiation} discusses the radiation characteristics of helical 
  tomotherapy, and \ref{comparisions} brielfy examines the differences between 
  this techique and other conformal therapy techniques. Section \ref{clinical} 
  is a collection of some clinical applications where tomotherapy 
  has, and in some cases has not,  been effective. Finally, section 
  \ref{conclusions} presents concluding 
  remarks on the topic.

@@

  \subsection{Construction}
  \label{construction}
  \begin{figure}
    \centering
    \subfloat[][]{
      \includegraphics[width=\linewidth]{fig5}
      \label{fig4a}}\\
    \subfloat[][]{
      \includegraphics[width=\linewidth]{fig4}
      \label{fig4b}}
    \caption{\protect\subref{fig4a} Axial rotation of the fan beam as the 
      patient (represented 
      as a cylinder) traverses laterally (Along the y-axis). 
      \protect\subref{fig4b} A simplified view of the Helical 
      Tomotherapy 
      treatment system (at a gantry angle of 0$^\circ$) in use today. The 
      setup consists of a 6 MV linac to produce an x-ray beam which is 
      collimated down to a fan-beam using 2 moveable jaws (denoted as `y-jaws' 
      in the image). There is also a Binary Multileaf to modulate the beam 
      laterally. \textit{Images courtesy of \cite{Fenwick2004}}}
    \label{fig4}
  \end{figure}
  Figure \ref{fig4} show the basics of dose delivery in the helical tomotherapy 
  system. There is a short in-line 6 MV linear accelerator that rotates on a 
  ring gantry at a source-axis of 85 cms (instead of the usual 100 cms). The 
  patient couch is translated through the gantry bore in the y-direction as 
  shown in \ref{fig4}\protect\subref{fig4a}. The pitch is about 0.2 to 0.5 
  (Pitch is the distance traveled by the couch in one rotation, divided by the 
  field width in the direction of traversal). The pair of y-jaws in 
  \ref{fig4}\protect\subref{fig4b} are used to define the y-width of the beam. 
  These are usually preset to values like 1,2.5, or 5 cm before imaging. 
  Lateral modulation of the beam is achieved using the binary multileaf. This 
  system consists of 64 binary multileaves, each having a width of 6.25 mm 
  projecting onto an isocenter. These leaves transition rapidly between on 
  and off states, the duration of which is made use of to modualate the 
  intensity. Other parameters are tabulated in \ref{tab1}.
  % Table generated by Excel2LaTeX from sheet 'Sheet1'
  \begin{table}[t]
    \centering
    \caption{Helical Tomotherapy Characteristics}
    \begin{tabularx}{1\linewidth}{XX}
      \toprule[0.3mm]
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      Gantry Rotation angle & 
      7\textsuperscript{$\circ$} \\
      Projections per revolution & 51 \\
      Rotation velocity & 10 - 60 sec/rotation \\
      Time-per projection & $>$ 196 ms \\
      \bottomrule
    \end{tabularx}%
    \label{tab1}%
  \end{table}%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             
  \begin{figure*}[htbp]
    \centering
    \subfloat[][]{
      \includegraphics[width=0.5 \linewidth]{fig6}
      \label{radchara}}
    \subfloat[][]{
      \includegraphics[width=0.5 \linewidth]{fig7}
      \label{radcharb}}
    \caption{\protect\subref{radchara} Shows the photon fluence profile of the 
    tomotherapy treatment beam. Notice the characteristic conical shape of the 
    tomotherapy profile (solid red line) and the flat shape of the conventional 
    linac IMRT system (dashed blue line). 
    \protect\subref{radcharb} Average photon beam spectra of the helical 
    tomography system in two operational modes: imaging (dashed green) and 
    treatment mode (solid red). \textit{Image courtesy of \cite{Jeraj2004}}}
    \label{fig5}
  \subsubsection{A note on Serial Tomotherapy}
  Serial tomography is another variant of tomotherapy that has been in use for 
  sometime now. The first clinical implementation occured as early as 1994. The 
  setup consists of a binary multileaf collimator attached to a conventional 
  linear accelerator. The collimator comprises of two opposing banks of 20 
  tungsten leaves which are driven pneumatically to lie either within or just 
  outside the fan beam radiation. The fan beam itself is modified by arranging 
  the leaves to lie within the radiation field for a variable amount of time.

  This can also be achieved by slightly modifying the conventional linear 
  accelerator by adding a binary multileaf collimator to the head of the linac. 
  Axial tomotherapy dose distributions are delivered slice-by-slice, with 
  patients being sequentially and discretely translated through the linac 
  gantry rotational plane between slices, helical distributions are delivered 
  without interruptions; patients are translated smoothly through the bore of 
  the machine as its gantry continuously and synchronously rotates, the 
  therapy equivalent of spiral computed tomography \cite{Fenwick2006}. 
  \subsection{Radiation Characteristics}
  \label{radiation}
  This section aims to briefly highlight the key differences in radiation 
  characteristics of helical tomography and other forms of radiotherapy. One of 
  the key differences is the lack of flattening filter, which makes the dose 
  more uniform are greater depths. As a result of this, the photon fluence 
  profile is shaped differently when compared to a traditional radiotherapy 
  system, as can be seen in Fig. \ref{radchara}. The conical shape of the 
  profile implies that there will be an increased average dose rate and thus 
  reduces the imaging time.

  Figure \ref{radcharb} shows the differences in the photon spectrum between 
  the 
  treatment and imaging modes of helical tomotherapy. Notice that the in the 
  imaging mode the spectrum has a lower energy when compared to treatment mode; 
  this is done so as to obatin better images by improving the contrast. A more 
  comprehensive comparision is documented in Ref. \cite{Jeraj2004}

  \subsection{Comparisions between IMRT techniques}
  \label{comparisions}
    \begin{table*}[t!]
      \centering
      \caption{A comparision between standard and non-standard IMRT delivery 
      systems (Ref. \cite{Fenwick2006})}
      \small{
      \begin{tabulary}{\linewidth}{LLLLLLL}
        \toprule[0.04cm]
        \textbf{System} & \textbf{Gantry} & \textbf{ Beam Geometry} & 
        \textbf{Collimated} & \textbf{Modulated} & \textbf{Imaging} & 
        \textbf{Gating
          Potential} \\\toprule[0.04cm]

@@

      \label{tab:compare}%
    \end{table*}%
  Tomotherapy is different from other standard IMRT techiniques such as 
  conformal radiotherapy (CRT) techniques and non-standard IMRT techniques like 
  Cyber knife.   Standard CRT techniques generate highly conformal radiotherapy 
  using radiation fields generated in a few fixed-angle positions. These are 
  generated by linear accelerators (aka. LINAC) which are then modulated by 
  conventional leaf collimators operated in either dynamic mode or the standard 
  step-and-shoot mode. CyberKnife is a modern radiotherapy system developed by 
  Accuray, Inc., primarily for radiosurgery. It makes use of a highly precise 
  robotic arm to delivery the doses. The arm is computer driven to achieve very 
  high precision, it is mounted with a compact 6-MV LINAC whose orientation is 
  determined by 6 different mechanical subsystems. This robotic arm can be 
  moved with a much greater generality than most conventional sytsems.

  Table \ref{tab:compare} highlights the key differences in various systems. 
  For the 
  sake the completeness, the authors of Ref. \cite{Fenwick2006} have 
  dicotomized the conventional system into 2 types, with and without a 
  multileaf collimator, and the tomotherapy into serial and parallel 
  tomotherapy.
  \section{Clinical Applications}
  \label{clinical}
  This section highlights few of the many clinical applications of tomotherapy. 
  \subsection{Total Body Irradiation (TBI)}
  Tomotherapy can be used in establishing total body irradiation (TBI) to gain 
  greater control over the dose distribution and to spare organs that may be at 
  risk.

  One of the disadvantages of using Helical Tomography (HT) for TBI is that it 
  can't be used when the body length of the patient exceeds 145 cms. For 
  patients who are taller than this, a better solution needs to be deviced.
  \subsection{Whole brain helical Tomotherapy}
  \subsection{Inoperable Lung Cancer}
  \subsection{Rectal Cancer}
  \section{Conclusion}
  \label{conclusions}
  \bibliography{Paper}{}
  \bibliographystyle{IEEEtran}
